---
title: "52 PROREF figures"
output: html_document
---

See 
`K:\Avdeling\Mar\NOG\Section 212\2018\Seksjonsm?te_23-24 april 2018\PROREF-publis`


```{r}
library(data.table)
library(dplyr)
library(purrr)
library(ggplot2)


# install.packages("xlsx")
# library(xlsx)

source("50fun_Calculate_PROREF_functions.R")
```

## Data
```{r}

# files have been copied from C:\Users\DHJ\OneDrive - NIVA\Documents\seksjon 212\Milkys 2017\Analyse\Data

# load(file = "Milkys_2017/Data/12d_data_all.RData")
load(file = "Input_data/2017/12d_data_all.RData")
# data_all from 12b4, this is the one to use for raw data

# load(file = "Milkys_2017/Data/12b3_data_all_2015sp.RData")        
load(file = "Input_data/2017/12b3_data_all_2015sp.RData")        
# data_all_2015sp - the better length-adjusted data
```


## Collect data - create data_all2
```{r}
# Unadjusted values
df1 <- df2 <- df3 <- data_all
df1$Basis <- "WW"
df1$Value <- df1$VALUE_WW
df1$Flag <- df1$FLAG1
df2$Basis <- "DW"
df2$Value <- df2$VALUE_DW
df2$Flag <- df2$FLAG1
df3$Basis <- "FB"
df3$Value <- df2$VALUE_FB
df3$Flag <- df3$FLAG1

# Length-adjusted values
df4 <- df5 <- df6 <- data_all_2015sp
df4$Basis <- "WWa"
df4$Value <- df4$VALUE_WWa
df4$Flag <- df4$FLAG1
df5$Basis <- "DWa"
df5$Value <- df5$VALUE_DWa
df5$Flag <- df5$FLAG1
df6$Basis <- "FBa"
df6$Value <- df6$VALUE_FBa
df6$Flag <- df6$FLAG1

# head(data_all, 1)
# head(data_all_2015sp, 1)

# dput(colnames(data_all_2015sp))
for (x in c("VALUE_WW", "VALUE_DW", "VALUE_FB", 
            "VALUE_WWa", "Flag_WWa", "VALUE_DWa", "VALUE_FBa", "Flag_DWa", "Flag_FBa", 
            "WTMEA", "SPECIMEN_NO", "SUBNO")
     ){
  df1[[x]] <- NULL
  df2[[x]] <- NULL
  df3[[x]] <- NULL
  df4[[x]] <- NULL
  df5[[x]] <- NULL
  df6[[x]] <- NULL
  }

if (mean(colnames(df1) == colnames(df4)) == 1){  # must be 1
  cat("Column names for 1 and 4 are identical\n")
} else {
  cat("Column names for 1 and 4 are NOT identical - must be fixed!\n")
}

# head(df1, 1)
# head(df4, 1)
# cbind(colnames(df1), colnames(df4))

data_all2 <- rbind(df1, df2, df3, df4, df5, df6)
nrow(data_all2) # 1922925

rm(df1, df2, df3, df4, df5, df6)

```

### Some corrections
```{r}
# Delete lines with no values
sel <- is.na(data_all2$Value); sum(sel)
data_all2 <- data_all2[!sel,]

# Some adjusted HBCDD values with no unit, set them to UG (why did this happen?)
sel <- with(data_all2, PARAM %in% "HBCDD")
data_all2$UNIT[sel] <- "UG_P_KG"

# TISSUE_NAME for blue mussel
sel <- with(data_all2, LATIN_NAME == "Mytilus edulis" & TISSUE_NAME != "Whole soft body")
data_all2$TISSUE_NAME[sel] <- "Whole soft body"

sel <- with(data_all2, YEAR < 1981); sum(sel)
data_all2 <- data_all2[!sel,]

nrow(data_all2) # 1272232

# sl? sammen  227G1 +  227G2 - already done
# xtabs(~STATION_CODE + YEAR + LATIN_NAME, subset(data_all2, substr(STATION_CODE,1,4) %in% "227G"))
```

```{r}
# Just for making R a little faster
# data_all2 <- subset(data_all2, Basis == "WW")

```

## Make variables for number (N_over_LOQ) and proportion (P_over_LOQ) of data above LOQ
```{r}
# Percentage of data above LOQ, by year and parameter
# tab <- xtabs(~YEAR + PARAM + is.na(Flag), data_all2)
# round(tab[,,"TRUE"]/apply(tab, 1:2, sum), 3)

data_all2 <- data_all2 %>%
  group_by(STATION_CODE, YEAR, LATIN_NAME, TISSUE_NAME, PARAM, Basis) %>%
  dplyr::mutate(N = n(), N_over_LOQ = sum(!grepl("<", FLAG1)))
data_all2$P_overLOQ <- with(data_all2, N_over_LOQ/N)

# Set to ordinary data frame
data_all2 <- as.data.frame(data_all2)

```

```{r}
years_backgr <- 2002:2016
```


## Example for paper 
### DDEPP 
```{r}
# plot_increase_determ("DDEPP", species = "Gadus morhua", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr))
# plot_increase("DDEPP", "Gadus morhua", "Lever", "VALUE_WW", min_years = 5, data = subset(data_all, YEAR %in% years_backgr))
par(mar = c(4,6,1,1))
plot_increase_alt("DDEPP", "Gadus morhua", "Lever", "VALUE_WW", min_years = 5, 
                  data = subset(data_all, YEAR %in% years_backgr))

```

### Same figure, saved to file
```{r}
# 
png("Figures/52_DDEPP_codliver.png", height = 12, width = 17, units = 'cm', res = 300, type="cairo", antialias="default")
par(mar = c(4,6,1,1))
plot_increase_alt("DDEPP", "Gadus morhua", "Lever", "VALUE_WW", min_years = 5, data = subset(data_all, YEAR %in% years_backgr))
dev.off()

```

### Tests - for adding manually to figure
```{r}
df <- differences_increasing_conc("DDEPP", sp = "Gadus morhua", "Lever", variable_name = "VALUE_WW",
                                  quantile = 0.9, data = subset(data_all, YEAR %in% years_backgr))

cbind(df$Test, round(df$P, 5))
```

### New version of plot for paper    
Version 2a, 2b    
```{r}

df <- plot_increase_data("DDEPP", "Gadus morhua", "Lever", "VALUE_WW", min_years = 5, 
                  data = subset(data_all, YEAR %in% years_backgr))

df <- df %>%
  mutate(Step = Rank - 1,
         GroupA = Test %>% stringr::str_extract("([^[[:blank:]]]+)"),
         GroupA_lineshift = gsub(",", "\n", GroupA)
  )
df <- df[-1,]

round_p <- function(x){
  case_when(
    x >= 0.01 ~ sprintf("P = %.2f", x),
    x >= 0.001 ~ sprintf("P = %.3f", x),
    TRUE ~ "P < 0.001")
}     
# round_p(0.345)
# round_p(0.075)
# round_p(0.004)
# round_p(0.0004)

# For the points
df2 <- df %>%
  select(Step, Median1, Median2) %>%
  tidyr::pivot_longer(-Step, names_to = "Data set", values_to = "Median") %>%
  mutate(`Data set` = case_when(
    `Data set` == "Median1" ~ "Data set 1",
    `Data set` == "Median2" ~ "Data set 2")
    )

gg <- ggplot(df2, aes(Step)) +
  geom_point(aes(y = Median, color = `Data set`), size = rel(3)) + 
  scale_color_manual("", values = c("blue", "red3")) + 
  geom_hline(yintercept = 0, size = 1) +
  geom_label(data = df, aes(y = Median2 + 20, label = Station), vjust = 0) +
  geom_label(data = df, aes(y = Median1 - 20, label = GroupA_lineshift), vjust = 1) +
  annotate("text", x = 1:7, y = -280, label = round_p(df$P)) +
  scale_x_continuous(breaks = 0:8, minor_breaks = seq(0.5, 8.5, 1)) +
  scale_y_continuous(breaks = c(0, 100, 200, 300), minor_breaks = seq(0, 350, 50)) +
  labs(x = "Step no.", y = "Median concentration") +
  coord_cartesian(ylim = c(-300, 390), xlim = c(0.4,7.8), expand = FALSE)

gg1 <- gg +
  theme(legend.position = c(0.02, 0.97),
        legend.justification = c("left", "top"),
        legend.margin = margin(0, 12, 5, 2),
        legend.title = element_blank(),
        panel.grid.major.x = element_blank())
gg2 <- gg +
  theme_bw() +
  theme(legend.position = c(0.02, 0.97),
        legend.justification = c("left", "top"),
        legend.margin = margin(0, 12, 5, 2),
        legend.title = element_blank(),
        panel.grid.major.x = element_blank())

# Original
ggsave("Figures/52_DDEPP_codliver_ver2a.png", gg1, 
       height = 12, width = 17, units = 'cm', dpi = 300, 
       type="cairo", antialias="default")

# Black and white
ggsave("Figures/52_DDEPP_codliver_ver2b.png", gg2, 
       height = 12, width = 17, units = 'cm', dpi = 300, 
       type="cairo", antialias="default")

gg2

```

### As above but showing only until step 4   
Version 2c, 2d  
```{r}
df_b <- df %>% filter(Step <= 4)
df2_b <- df2 %>% filter(Step <= 4)

gg <- ggplot(df2_b, aes(Step)) +
  geom_point(aes(y = Median, color = `Data set`), size = rel(3)) + 
  scale_color_manual("", values = c("blue", "red3")) + 
  # geom_hline(yintercept = 0, size = 1) +
  geom_label(data = df_b, aes(y = Median2 + 4, label = Station), vjust = 0, color = "red3") +
  geom_label(data = df_b, aes(y = Median1 - 4, label = GroupA_lineshift), vjust = 1, color = "blue") +
  annotate("text", x = 1:4, y = 2, vjust = 0,
           label = paste0("Difference:\n", round_p(df_b$P))) +
  scale_x_continuous(breaks = 0:4, minor_breaks = seq(0.5, 4.5, 1)) +
  scale_y_continuous(breaks = seq(0, 100, 25), minor_breaks = seq(0, 150, 12.5)) +
  labs(x = "Step no.", y = expression(Median~concentration~(mu*g/kg))) +
  coord_cartesian(ylim = c(0, 60), xlim = c(0.4,4.8), expand = FALSE)

gg1 <- gg +
  theme(legend.position = c(0.05, 0.97),
        legend.justification = c("left", "top"),
        legend.margin = margin(0, 12, 5, 2),
        legend.title = element_blank(),
        legend.text = element_text(size = rel(1.2)),
        axis.text.x = element_text(size = rel(1.4)),
        panel.grid.major.x = element_blank())
gg2 <- gg +
  theme_bw() +
  theme(legend.position = c(0.05, 0.97),
        legend.justification = c("left", "top"),
        legend.margin = margin(0, 12, 5, 2),
        legend.title = element_blank(),
        legend.text = element_text(size = rel(1.2)),
        axis.text.x = element_text(size = rel(1.4)),
        panel.grid.major.x = element_blank())

# Original
ggsave("Figures/52_DDEPP_codliver_ver2c.png", gg1,
       height = 12, width = 17, units = 'cm', dpi = 300,
       type="cairo", antialias="default")

# Black and white
ggsave("Figures/52_DDEPP_codliver_ver2d.png", gg2,
       height = 12, width = 17, units = 'cm', dpi = 300,
       type="cairo", antialias="default")

gg2

```

## Process, step by step
### Step 1: get data series (for several stations). Result: object with medians (per year) and medians (per station)
```{r}
# debugonce(get_lower_medians)
X <- get_lower_medians("DDEPP", sp = "Gadus morhua", ti = "Lever", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr))
str(X, 1)
str(X$data_medians_list[["10B"]], 1)
X$median_per_station
```

### find_set_differences
```{r}
diff <- find_set_differences(X)
diff
```

### Plot first station 
```{r}
i <- 1
diff$Station[i]
ggplot(X$data_medians_list[[diff$Station[i]]], aes(x = STATION_CODE, y = median)) + geom_point()

# ggplot(X$data_medians_list[diff$Station[i:(i+1)]], aes(x = STATION_CODE, y = median)) + geom_point()
```

### Plot first vs. second station 
```{r}
# debugonce(find_set_difference)
# diff <- find_set_difference(X, 1)
diff <- find_set_difference(X, 2)

```

### Making a sorted station list
```{r}
############################################################################################
#
# Example of making a sorted station list (from least to most contaminanted) and make a "background station" plot 
# Note: slightly outdated, the procedure now is a bit different
#
############################################################################################

#
# Step 1: get data series (for several stations). Result: object with medians (per year) and medians (per station)
#
years_backgr <- 2002:2016
# debugonce(get_lower_medians)
X <- get_lower_medians("NI", sp = "Gadus morhua", ti = "Lever", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr))
str(X, 1)
# $ data_medians_list :List of 11
# $ median_per_station: Named num [1:11] 0.05 0.11 0.045 0.045 0.1 0.07 0.046 0.059 0.0525 0.055 ...
#  ..- attr(*, "names")= chr [1:11] "10B" "13B" "15B" "23B" ...



#
# Step 2:Find difference  between (example): station 1-3 vs. station 4, when stations are ordered 
#   in increasing order by median
#
debugonce(find_set_difference)
find_set_difference(X, 4)

#
# Step 3: Find all differences 1 vs. 2, 1-2 vs. 3, 1-3 vs. 4, etc. (stations in increasing order)
#
find_set_differences(X)
# This does the same but returns NULL if there is no data 
differences_increasing_conc("NI", sp = "Gadus morhua", ti = "Lever", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr))

#
# Step 4: Make plot of differences among stations in increasing order for one compound, species and tissue
#
plot_increase_determ("NI", species = "Gadus morhua", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr))

#
# Step 5: Make all plots for a group of chemicals; don't have to specify tissue; plot to file by default
#
determ <- c("CD", "PB", "HG", "AS", "CR", "ZN", "NI", "CO", "CU", "AG")
plot_increase_mult(determ, species = "Gadus morhua", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr), 
                   determinantgroup = "metals", plot = "window")


```


### Creating median data

#### Turn data into data table and check that there is only one UNIT per series
```{r}
data_all2_dt <- setDT(data_all2)

check_unit <- data_all2_dt[!is.na(Value), .(n_unit = length(unique(UNIT))),
                           by = .(STATION_CODE, TISSUE_NAME, LATIN_NAME, PARAM, YEAR, Basis)]
table(check_unit$n_unit)


```

## Calculate medians using data.table() 
Divided in two parts
```{r}
t0 <- Sys.time()

data_all_median1 <- 
  data_all2_dt[!is.na(Value),
               .(N = .N, median(Value), SD = sd(Value), Unit = head(UNIT, 1), Over_LOQ = sum(is.na(Flag))),
               by = .(STATION_CODE, TISSUE_NAME, LATIN_NAME, PARAM, YEAR, Basis)]

data_all_median2 <- 
  data_all2_dt[!is.na(Flag) & !is.na(Value),
               .(Det_limit = median(Value), OverLOQ_min = min(Value), OverLOQ_max = max(Value)),
               by = .(STATION_CODE, TISSUE_NAME, LATIN_NAME, PARAM, YEAR, Basis)]

data_all_median <- left_join(data_all_median1, data_all_median2, 
                             by = c("STATION_CODE", "TISSUE_NAME", "LATIN_NAME", "PARAM", "YEAR", "Basis"))

t0 - Sys.time()
```






### Calculate medians
```{r}
# Put calculation of 'Value' last because otherwise the wrong 'Value' is used for Det_limit etc.

# 40 seconds
t0 <- Sys.time()
data_all_median <- data_all2 %>%
  group_by(STATION_CODE, TISSUE_NAME, LATIN_NAME, PARAM, YEAR, Basis) %>%
  summarise(N = n(), 
            Over_LOQ = sum(is.na(Flag)), 
            Det_limit = ifelse(Over_LOQ < N, median(Value[!is.na(Flag)]), NA), 
            OverLOQ_min = ifelse(Over_LOQ < N, min(Value[!is.na(Flag)]), NA), 
            OverLOQ_max = ifelse(Over_LOQ < N, max(Value[!is.na(Flag)]), NA), 
            SD = sd(Value, na.rm = TRUE),
            Unit = first(UNIT), Unit_n = n_distinct(UNIT), 
            Value = median(Value))
data_all_median$P_over_LOQ <- with(data_all_median, Over_LOQ/N)
t0 - Sys.time()

# Medians for times series analysis: Set less than LOQ-data to 50%
rawdata_for_regression <- data_all2
sel <- !is.na(rawdata_for_regression$FLAG1)
mean(sel)
rawdata_for_regression$Value[sel] <- 0.5*rawdata_for_regression$Value[sel]

# 40 seconds
t0 <- Sys.time()
data_all_median_for_regression <- rawdata_for_regression %>%
  group_by(STATION_CODE, TISSUE_NAME, LATIN_NAME, PARAM, YEAR, Basis) %>%
  summarise(N = n(), 
            Over_LOQ = sum(is.na(Flag)), 
            Det_limit = ifelse(Over_LOQ < N, median(Value[!is.na(Flag)]), NA), 
            OverLOQ_min = ifelse(Over_LOQ < N, min(Value[!is.na(Flag)]), NA), 
            OverLOQ_max = ifelse(Over_LOQ < N, max(Value[!is.na(Flag)]), NA), 
            SD = sd(Value, na.rm = TRUE),
            Unit = first(UNIT), Unit_n = n_distinct(UNIT), 
            Value = median(Value))
data_all_median_for_regression$P_over_LOQ <- with(data_all_median_for_regression, Over_LOQ/N)
t0 - Sys.time()

xtabs(~Unit_n, data_all_median)                 # Only 1 as it should be (174848 rows)
xtabs(~Unit_n, data_all_median_for_regression)  # Only 1 as it should be (174848 rows)

```


