---
title: "53_Proref_revision_2024"
format: html
---

## Packages, loading functions

```{r}
library(data.table)
library(dplyr)
library(purrr)
library(ggplot2)
library(forcats)
library(ggeasy)
library(tidyr)
library(ggrepel)

# install.packages("xlsx")
# library(xlsx)

source("50fun_Calculate_PROREF_functions.R")

df_old_result <- readxl::read_excel("data/Proref_2022_version1.xlsx")
df_stations <- readxl::read_excel("data/Kartbase_edit.xlsx")

# For printout:
hyphens <- "-------------------------------------------------------------------------------\n"

```


## Data
```{r}

# files have been copied from C:\Users\DHJ\OneDrive - NIVA\Documents\seksjon 212\Milkys 2017\Analyse\Data

# load(file = "Milkys_2017/Data/12d_data_all.RData")
load(file = "Input_data/2017/12d_data_all.RData")
# data_all from 12b4, this is the one to use for raw data

# load(file = "Milkys_2017/Data/12b3_data_all_2015sp.RData")        
load(file = "Input_data/2017/12b3_data_all_2015sp.RData")        
# data_all_2015sp - the better length-adjusted data
```



### Some corrections
```{r}

# Delete lines with no values
sel <- is.na(data_all$VALUE_WW); sum(sel)
data_all <- data_all[!sel,]

# Some adjusted HBCDD values with no unit, set them to UG (why did this happen?)
sel <- with(data_all, PARAM %in% "HBCDD")
# table(addNA(data_all$UNIT[sel]))
# data_all$UNIT[sel] <- "UG_P_KG"

# TISSUE_NAME for blue mussel
sel <- with(data_all, LATIN_NAME == "Mytilus edulis" & TISSUE_NAME != "Whole soft body")
data_all$TISSUE_NAME[sel] <- "Whole soft body"

sel <- with(data_all, YEAR < 1981); sum(sel)
data_all <- data_all[!sel,]

nrow(data_all) # 319179

# sl? sammen  227G1 +  227G2 - already done
# xtabs(~STATION_CODE + YEAR + LATIN_NAME, subset(data_all, substr(STATION_CODE,1,4) %in% "227G"))
```


### Define years  
```{r}
years_backgr <- 2002:2016
```


## Process, step by step
### Step 1: get_lower_medians  

- get data series for one substance and all stations  
- Result: object with medians (per year) and medians (per station)  

```{r}
# debugonce(get_lower_medians)
X <- get_lower_medians("DDEPP", sp = "Gadus morhua", ti = "Lever", variable = "VALUE_WW", 
                       data = subset(data_all, YEAR %in% years_backgr))
# X <- get_lower_medians("PFOSA", sp = "Gadus morhua", ti = "Lever", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr))
cat(hyphens, "Result object \n", hyphens)
str(X, 1)
cat(hyphens, "'data_medians_list' for '10B'\n", hyphens)
str(X$data_medians_list[["10B"]])
cat(hyphens, "'median_per_station' (here, sorted)\n", hyphens)
sort(X$median_per_station)

```
### find_set_difference  
```{r}

i <- 3

# debugonce(find_set_difference)
a_single_test <- find_set_difference(X, i)

cat(hyphens, "Result object, testing station no. 1 -", i-1, "vs. station no.", i, "\n", hyphens)
str(a_single_test)

```


### find_set_differences  

- Find all differences 1 vs. 2, 1-2 vs. 3, 1-3 vs. 4, etc. (stations in increasing order)  

```{r}

# debugonce(find_set_differences)
diff <- find_set_differences(X) 

```

### differences_increasing_conc  

- 'differences_increasing_conc' is just 'get_lower_medians' followed by 'find_set_differences'  
- this function is used in 'plot_increase'  

```{r}

diff <- differences_increasing_conc(
  "DDEPP", sp = "Gadus morhua", ti = "Lever", variable = "VALUE_WW", 
  data = subset(data_all, YEAR %in% years_backgr))

# differences_increasing_conc(
#   "PFOSA", sp = "Gadus morhua", ti = "Lever", variable = "VALUE_WW", 
#   data = subset(data_all, YEAR %in% years_backgr))

```

### Plot first station   

- with percent over LOQ shown  

```{r}

i <- 1
diff$Station[i]
ggplot(X$data_medians_list[[diff$Station[i]]] %>% mutate(perc_over_loq = round(100*n_loq/n,0)), 
       aes(x = STATION_CODE, y = median)) + 
  geom_point() +
  geom_text(aes(label = perc_over_loq), hjust = 0, nudge_x = 0.05)

# ggplot(X$data_medians_list[diff$Station[i:(i+1)]], aes(x = STATION_CODE, y = median)) + geom_point()
```
### Plot all stations  

- with percent over LOQ shown  

```{r}

ggplot(diff %>% mutate(Station = fct_inorder(factor(Station))), 
       aes(Station, Median2)) +
  geom_point() +
  geom_text(aes(label = Perc_over_loq2), hjust = 0, nudge_x = 0.1)

```
### Plot all stations with 'median ratio'   

- 'median ratio' is the ratio between the median of a station and the median of the first station  

```{r}

ggplot(diff %>% mutate(Station = fct_inorder(factor(Station))), 
       aes(Station, Median2)) +
  geom_point() +
  geom_text(aes(label = round(Median_ratio, 1)), hjust = 0, nudge_x = 0.1)

```

### plot_increase  

* 'plot_increase' is 'differences_increasing_conc' followed by plotting  
* returns the rsult of 'differences_increasing_conc'  

```{r}

# DDEPP", sp = "Gadus morhua", ti = "Lever", variable = "VALUE_WW"

diff <- plot_increase(par = "DDEPP", sp = "Gadus morhua", ti = "Lever", variable_name = "VALUE_WW",
              data = subset(data_all, YEAR %in% years_backgr))  

diff <- plot_increase(par = "CB28", sp = "Mytilus edulis", ti = "Whole soft body", variable_name = "VALUE_WW",
                      min_indiv_per_year = 2, 
                      data = subset(data_all, YEAR %in% years_backgr))  


# debugonce(plot_increase)
# debugonce(differences_increasing_conc)
# debugonce(get_lower_medians)

```

### get_background_values 

- like 'differences_increasing_conc', 'get_background_values' is 'get_lower_medians' 
followed by 'find_set_differences'  
- but before that it also sets 'min_indiv_per_year' and tissue based on determinant (substance) and species   
```{r}

# debugonce(get_background_values)
one_case <- get_background_values(determinant = "DDEPP", species = "Gadus morhua", var_name = "VALUE_WW", 
                              years_backgr = 1992:2016,
                              data = subset(data_all, YEAR %in% years_backgr))  

str(one_case, 1)

```

### get all selected background values   



```{r}

# "Safe" version of function (doesn't break the loop if there are errors)
get_background_values_s <- safely(get_background_values)

df_series_sel <- readxl::read_excel("Info/Proref table - MS version 17.xlsx")

cat(hyphens, "Number of selected compounds (per species)\n", hyphens)  
df_series_sel$Species %>% table()

result_list <- map2(
  df_series_sel$PARAM, 
  df_series_sel$Species,
     \(x,y) get_background_values_s(determinant = x, species = y, var_name = "VALUE_WW", 
                              years_backgr = 1992:2016,
                              data = subset(data_all, YEAR %in% years_backgr))
     )

ok <- map(result_list, "error") %>% map_lgl(is.null)
cat(hyphens, "Number of successful compound*species\n", hyphens)
cat(sum(ok), "out of", length(ok), "\n")

result <- map_dfr(result_list[ok], c("result", "result_one_line"))  
cat(hyphens, "Result\n", hyphens)
str(result)

```

## Data through 2022  

### Raw data for analysis
```{r}
data_all2 <- readRDS("Input_data/2022/105_data_with_uncertainty_2023-09-12.rds") %>%
  rename(YEAR = MYEAR)
```

#### Remove some stations  
```{r}

# "B" mussel stations (industry data, from Langøya?)
xtabs(~STATION_CODE, data_all2 %>% filter(substr(STATION_CODE, 1, 1) == "B"))

data_all2 <- data_all2 %>%
  filter(!substr(STATION_CODE, 1, 1) == "B")

# Atations used for very few years
check <- data_all2 %>%
  distinct(STATION_CODE, YEAR) %>%
  count(STATION_CODE) %>%
  arrange(n)
# check
table(check$n)

# Remove those with 4 years and less
data_all2 <- data_all2 %>%
  group_by(STATION_CODE) %>%
  mutate(no_years = length(unique(YEAR))) %>%
  filter(no_years >= 5)

check <- data_all2 %>%
  distinct(STATION_CODE, YEAR) %>%
  count(STATION_CODE) %>%
  arrange(n)
# check
table(check$n)


```


#### Change some parameters in list of parameters to be analysed  

```{r}

xtabs(~PARAM, data_all2 %>% filter(grepl("SCCP", PARAM)))
xtabs(~PARAM, df_series_sel %>% filter(grepl("SCCP", PARAM)))

xtabs(~PARAM, data_all2 %>% filter(grepl("PAH", PARAM)))
xtabs(~PARAM, df_series_sel %>% filter(grepl("PAH", PARAM)))

df_series_sel <- df_series_sel %>%
  mutate(
    PARAM = case_when(
      grepl("SCCP", PARAM) ~ "SCCP eksl. LOQ",
      grepl("MCCP", PARAM) ~ "MCCP eksl. LOQ",
      grepl("PAH", PARAM) ~ "PAH16_exloq",
      TRUE ~ PARAM
    )
  )

```
### Run analysis for data through 2022  

```{r}

cat(hyphens, "Number of selected compounds (per species)\n", hyphens)  
df_series_sel$Species %>% table()

years_backgr2 <- 1992:2022

result_list <- map2(df_series_sel$PARAM, 
     df_series_sel$Species,
     \(x,y) get_background_values_s(determinant = x, species = y, var_name = "VALUE_WW", 
                              years_backgr = 1992:2022,
                              data = subset(data_all2, YEAR %in% years_backgr2))
     )

ok <- map(result_list, "error") %>% map_lgl(is.null)
cat(hyphens, "Number of successful compound*species\n", hyphens)
cat(sum(ok), "out of", length(ok), "\n")

result <- map_dfr(result_list[ok], c("result", "result_one_line"))  
cat(hyphens, "Result\n", hyphens)
str(result)

writexl::write_xlsx(result, "Data/Proref_2022_version1.xlsx")   

```

### Get detailed results  

```{r}

# str(result_list[[1]]$result, 1)

result_detailed <- map_dfr(result_list[ok], c("result", "differences"))  
cat(hyphens, "Detailed result\n", hyphens)
str(result_detailed)

threshold_p1 <- 0.10
threshold_p2 <- 0.05
threshold_medianratio <- 2

# Adding columns for background and last rank  

result_detailed <- result_detailed %>% 
    # filter(PARAM %in% "CU" & LATIN_NAME %in% species) %>%   
    group_by(PARAM, LATIN_NAME, TISSUE_NAME) %>%
    mutate(
      P_round = round(P, 3),       # just for more easily readable output
      P_prev = lag(P_round),        # previous P value
      P_change1 = case_when(
        Rank == 1 ~ FALSE,
        TRUE ~ (P < threshold_p1) & (P_prev >= threshold_p1)), # change in P value
      P_cum1 = cumsum(P_change1), # cumulative change
      P_change2 = case_when(
        Rank == 1 ~ FALSE,
        TRUE ~ (P < threshold_p2) & (P_prev >= threshold_p2)), # change in P value
      P_cum2 = cumsum(P_change2), # cumulative change
      Background1 = case_when(
        P_cum1 == 0 ~ "Background",
        TRUE ~ "Other"),
      Background1b = case_when(
        P_cum1 == 0 & Median_ratio <= threshold_medianratio ~ "Background",
        TRUE ~ "Other"),
      Background2 = case_when(
        P_cum2 == 0 ~ "Background",
        TRUE ~ "Other"),
      Background = case_when(
        Background1b == "Background" ~ "Background 1b and 2",
        Background1 == "Background" ~ "Background 1 and 2",
        Background2 == "Background" ~ "Background 2",
        TRUE ~ "Other"),
      # This is used only for adding PARAM label to the last station
      #   in plots:
      Last_rank = (Rank == max(Rank))
      ) %>% 
  ungroup() %>%
    mutate(
      Line_label = case_when(
        Last_rank ~ PARAM,
        TRUE ~ NA)
    ) #%>%
  #select(-P_prev, -P_change1, -P_cum1, -P_change2, -P_cum2, -Last_rank)

```

#### Table check 

```{r}

xtabs(~Background, result_detailed)

```


#### Plotting 'result_detailed'  

Example: metals in cod  

```{r}

determ <- c("CD", "PB", "HG", "AS", "CR", "ZN", "NI", "CO", "CU", "AG")
determ <- c("SCCP eksl. LOQ", "MCCP eksl. LOQ")
species <- "Gadus morhua"

pointcolors <- c("Background 1b and 2" = "blue4", 
               "Background 1 and 2" = "purple", 
               "Background 2" = "green3", 
               "Other" = "orange")

gg <- ggplot(
  result_detailed %>% filter(PARAM %in% determ & LATIN_NAME %in% species), 
  aes(x = Rank, y = Median2, group = PARAM)) +
  geom_line() +  
  geom_point(aes(fill = Background), pch = 21, size = rel(2)) +
  scale_fill_manual(values = pointcolors) +
  geom_text(aes(label = Line_label), hjust = 0, nudge_x = 0.35, size = rel(4))+
  labs(title = "Concentrations", x = "Station ranking", y = "Concentration")

# gg
gg + scale_y_log10()



ggplot(
  result_detailed %>% filter(PARAM %in% determ & LATIN_NAME %in% species), 
  aes(x = Rank, y = Median_ratio, group = PARAM)) +
  geom_line() +  
  geom_point(aes(fill = Background), pch = 21, size = rel(2)) +
  scale_fill_manual(values = pointcolors) +
  geom_text(aes(label = Line_label), hjust = 0, nudge_x = 0.35, size = rel(4))+
  labs(title = "Concentrations ratio to first station", 
       x = "Station ranking", y = "Concentration ratio")


```

#### Plot 'median ratio'

- note: not including 'Background 1b' points  
    - this is because the 'Background 1b' 

```{r}

# 
resultlist <- list(
    result_detailed %>% 
      filter(Background1 %in% "Background" & LATIN_NAME == "Gadus morhua") %>%
      mutate(Background = "Background 1"),
    result_detailed %>% 
      filter(Background2 %in% "Background" & LATIN_NAME == "Gadus morhua") %>%
      mutate(Background = "Background 2"),
    result_detailed %>% 
      filter(Background1 %in% "Background" & LATIN_NAME == "Mytilus edulis") %>%
      mutate(Background = "Background 1"),
    result_detailed %>% 
      filter(Background2 %in% "Background" & LATIN_NAME == "Mytilus edulis") %>%
      mutate(Backg = "Background 2"))

result_detailed_summ_all <- map_dfr(
  resultlist,
  \(df) df  %>%
    group_by(PARAM, LATIN_NAME, TISSUE_NAME, UNIT) %>%
    summarize(
      Background = first(Background),
      Stations = paste(Station, collapse = ", "),
      Median_ratio = max(Median_ratio),
      n = n(), .groups = "drop"
    ))

gg <- map(
  c("Gadus morhua", "Mytilus edulis"),
  \(species)
  ggplot(result_detailed_summ_all %>% filter(LATIN_NAME == species),
         aes(x = n, y = Median_ratio, color = Background)) +
  geom_point() + 
  geom_text_repel(aes(label = PARAM), hjust = 0, nudge_x = 0.2, size = rel(3),
                  max.overlaps = 5) +
  labs(title = paste("Median ratio -", species),
       x = "Number of background stations", y = "Max median ratio")
)

gg[[1]]
# gg[[1]] + scale_y_log10()
gg[[2]]
    
```


```{r}

resultlist <- list(
    result_detailed %>% 
      filter(Background1b %in% "Background") %>%
      mutate(Background = "Background 1b"),
    result_detailed %>% 
      filter(Background1 %in% "Background") %>%
      mutate(Background = "Background 1"),
    result_detailed %>% 
      filter(Background2 %in% "Background") %>%
      mutate(Background = "Background 2"))

result_summ <- map_dfr(
  resultlist,
  \(df) df  %>%
    group_by(PARAM, LATIN_NAME) %>%
    summarize(
      Background = first(Background),
      Stations = paste(Station, collapse = ", "),
      Median_ratio = max(Median_ratio),
      n = n(), .groups = "drop"
    ))  %>%
  arrange(PARAM, LATIN_NAME, Background) 

# View(result_summ)

```


### 1b

```{r}

# df_old_result %>%
#   filter(LATIN_NAME == "Gadus morhua") %>% pull(PARAM)  %>% dput()
# df_old_result %>%
#   filter(LATIN_NAME == "Mytilus edulis") %>% pull(PARAM) %>% dput()

param_levels <- c("HG", "CD", "CR", "CU", "NI", "PB", "AG", "ZN", "AS", 
                  "CB_S7", "DDEPP", "HCHG", "HCB", 
                  "NAP", "ANT", "FLU", "BAA", "BAP", "PAH16_exloq", 
                  "4-N-NP", "4-N-OP", "4-T-NP", "4-T-OP", 
                  "BDE47", "BDE6S", "HBCDA", "PFOA", "PFOS", "PFOSA", 
                  "SCCP eksl. LOQ", "MCCP eksl. LOQ", "TBT")


result_1b <- result_detailed %>% 
  filter(!is.na(Background1b) & PARAM %in% df_old_result$PARAM) %>%
  left_join(df_stations, by = c("Station" = "STATION_CODE"), relationship = "many-to-one") %>%
  arrange(desc(Station_order)) %>%
  mutate(
    Station_name = fct_inorder(paste(Station, Station_short, sep = "_")), 
    PARAM = factor(PARAM, levels = param_levels)) 

```
#### Plot cod  
```{r}

result_1b %>%
  filter(LATIN_NAME == "Gadus morhua") %>% 
  ggplot(aes(x = PARAM, y = Station_name, fill = Background1b)) +
  geom_tile() +
  scale_fill_manual(values = c("Background" = "blue4", "Other" = "orange")) +
  easy_rotate_x_labels(angle = -45) +
  labs(title = "Background 1b, cod", x = "Parameter", y = "Station")

```

#### Plot blue mussel

```{r}

#| fig-width: 6
#| fig-height: 12

gg <- result_1b %>%
  filter(LATIN_NAME == "Mytilus edulis") %>% 
  ggplot(aes(x = PARAM, y = Station_name, fill = Background1b)) +
  geom_tile() +
  scale_fill_manual(values = c("Background" = "blue4", "Other" = "orange")) +
  easy_rotate_x_labels(angle = -45) +
  labs(title = "Background 1b, blue mussel", x = "Parameter", y = "Station")

gg

```
#### Stations  
```{r}

data_all2 %>%
  distinct(STATION_CODE, YEAR) %>%
  group_by(STATION_CODE) %>%
  summarize(n = n(), last_year = max(YEAR), .groups = "drop")

```

In 2007, the mussels were only insignificantly contaminated (Class I) except for the moderate 
concentrations (Class II) found at six stations remote from point sources (43A Lyngneset in the 
Langfjord (Arnøy, northwest of Skjervøy), 45A Sauhamneset in Revsbotn (Sørøysundet), 47A 
Kifjordneset on the southwest coast of Nordkyn peninsula (Laksefjorden), 49A Norfjorden 
(Syltefjord), 10A2 Skallneset (north side of Varangerfjord), and 11X Brashavn (south side of 
Varangerfjord)). Five of these cases were due to cadmium and indicate a natural regional difference 
(Appendix I and Appendix J). T

```{r}

dput(sort(unique(result_1b$PARAM)))

c("AG", "ANT", "AS", "CD", "CR", "CU", "HG", "NI", "PB", "ZN".
  

"BAA", "BAP", "BDE47", "BDE6S", "BDESS", "CB_S7", "DDEPP", "FLU", "HBCDA", "HCB", "HCHG",, "MCCP eksl. LOQ", 
"NAP", "PAH16_exloq", "PFOA", "PFOS", "PFOSA", "SCCP eksl. LOQ", 
"TBT", "4-N-NP", "4-N-OP", "4-T-NP", "4-T-OP", )




result_1b_for_matrix <- result_detailed %>% 
  filter(!is.na(Background1b)) %>%
  mutate(Param_species = paste(PARAM, LATIN_NAME, sep = "_")) %>%
  distinct(Param_species, Station, Background1b)
# apply(!is.na(result_1b_natrix), 2, sum)

result_1b_for_matrix <- result_detailed %>% 
  filter(!is.na(Background1b)) %>%
  filter(LATIN_NAME == "Gadus morhua") %>%
  ggplot(aes(x = PARAM, y = Station, fill = Background1b)) +
  geom_tile() +
  easy_rotate_x_labels(angle = -45)
  


```


#### Plot matrix cod  
```{r}

result_detailed %>%
  filter(LATIN_NAME == "Gadus morhua") %>%
  count(PARAM) %>%
  arrange(desc(n))
result_detailed %>%
  filter(LATIN_NAME == "Gadus morhua") %>%
  count(Station) %>%
  arrange(desc(n))

result_detailed %>%
  filter(LATIN_NAME == "Gadus morhua") %>%
  add_count(PARAM, name = "count_param") %>%
  add_count(Station, name = "count_station") %>%
  filter(count_param >= 10) %>%
  # filter(count_param >= 10 & count_station >= 11) %>%
  ggplot(aes(x = PARAM, y = Station, fill = Background)) +
  geom_tile() +
  easy_rotate_x_labels(angle = -45)

```


```{r}

result_matrix_all <- result_detailed %>%
  mutate(Param_species = paste(PARAM, LATIN_NAME, sep = "_")) %>%
  count(Param_species, Station) %>%
  pivot_wider(names_from = Param_species, values_from = n, values_fill = 0)



result_detailed %>%
  mutate(Param_species = paste(PARAM, LATIN_NAME, sep = "_")) %>%
  count(Param_species) %>%
  arrange(desc(n))
result_detailed %>%
  count(Station) %>%
  arrange(desc(n))

result_detailed %>%
  mutate(Param_species = paste(PARAM, LATIN_NAME, sep = "_")) %>%
  add_count(Param_species, name = "count_param") %>%
  add_count(Station, name = "count_station") %>%
  filter(count_param >= 15 & count_station >= 10) %>%
  ggplot(aes(x = Param_species, y = Station, fill = Background)) +
  geom_tile() +
  easy_rotate_x_labels(angle = -45)
  
```


### Get summary result   

#### original function  
```{r}

result_summ <- map_dfr(result_list[ok], c("result", "result_one_line"))  

# str(result_list[[1]], 1)
# str(result_list[[1]]$result, 1)

```





```{r}

X$median_per_station

```


### Plot first vs. second station 
```{r}
# debugonce(find_set_difference)
# diff <- find_set_difference(X, 1)
# diff <- find_set_difference(X, 2)

i <- 1:2
ggplot(X$data_medians_list[[diff$Station[i]]], aes(x = STATION_CODE, y = median)) + geom_point()

```

### Making a sorted station list
```{r}
############################################################################################
#
# Example of making a sorted station list (from least to most contaminanted) and make a "background station" plot 
# Note: slightly outdated, the procedure now is a bit different
#
############################################################################################

#
# Step 1: get data series (for several stations). Result: object with medians (per year) and medians (per station)
#
years_backgr <- 2002:2016
# debugonce(get_lower_medians)
X <- get_lower_medians("NI", sp = "Gadus morhua", ti = "Lever", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr))
str(X, 1)
# $ data_medians_list :List of 11
# $ median_per_station: Named num [1:11] 0.05 0.11 0.045 0.045 0.1 0.07 0.046 0.059 0.0525 0.055 ...
#  ..- attr(*, "names")= chr [1:11] "10B" "13B" "15B" "23B" ...



#
# Step 2:Find difference  between (example): station 1-3 vs. station 4, when stations are ordered 
#   in increasing order by median
#
debugonce(find_set_difference)
find_set_difference(X, 4)

#
# Step 3: Find all differences 1 vs. 2, 1-2 vs. 3, 1-3 vs. 4, etc. (stations in increasing order)
#
find_set_differences(X)
# This does the same but returns NULL if there is no data 
differences_increasing_conc("NI", sp = "Gadus morhua", ti = "Lever", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr))

#
# Step 4: Make plot of differences among stations in increasing order for one compound, species and tissue
#
plot_increase_determ("NI", species = "Gadus morhua", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr))

#
# Step 5: Make all plots for a group of chemicals; don't have to specify tissue; plot to file by default
#
determ <- c("CD", "PB", "HG", "AS", "CR", "ZN", "NI", "CO", "CU", "AG")
plot_increase_mult(determ, species = "Gadus morhua", variable = "VALUE_WW", data = subset(data_all, YEAR %in% years_backgr), 
                   determinantgroup = "metals", plot = "window")


```

