---
title: "53_Proref_revision_2024"
format: html
---

This script runs the PROREF procedure for all stations, parameters and basises  
- Three "stopping" rules are used (Background1, Background1b and Background2)  
- The final results are for Background1 (all stations not being different at P < 0.10, and the ratio of concentration medians between the cleanest and dirties background station is not allowed to be above 2  
- The whole script is a bit slow - it takes about 1 hour 15 minutes 
    - 3.2 Select background stations: Run analysis for all data: 22 min.
    - (one hour for "Run analysis for data through 2022", one hour for "add the actual proref")  

## 1. Startup

### Packages, loading functions  

```{r}
library(data.table)
library(dplyr)
library(purrr)
library(ggplot2)
library(forcats)
library(ggeasy)
library(tidyr)
library(glue)
library(ggrepel)
library(stringr)
# install.packages("furrr")
library(furrr)

# install.packages("xlsx")
# library(xlsx)

source("54fun_Calculate_PROREF_functions.R")
source("54data_Proref_2024_functions.R")       # for get_coverage and get_coverage_species 

# For parallell processing (coding not finished)
# - see "Run analysis for data through 2022, parallel processing"
# library(doParallel)
# cl <- makeCluster(10)
# registerDoParallel(cores = 10)

years_backgr2 <- 2003:2022

# For printout:
hyphens <- "-------------------------------------------------------------------------------\n"

```

### Settings  

```{r}

# reruns analysis of 3.2   
# options 
rerun_analysis_values <- c("all", "no", "test")
# "all":   reruns entire analysis (25 minutes with 4 cores)
# "no":    uses saved results, given by 'saved_detailed_results'   
# "test":  reruns analysis for a mini test (two series only)  

rerun_analysis <- "test"
save_results <- FALSE
saved_detailed_results <- "Data/54_result_detailed_2003-2022_2025-01-20-T1603.rds"

if (!(rerun_analysis %in% rerun_analysis_values)){
  stop("'rerun_analysis' must be one of these: ", paste(rerun_analysis_values, collapse = ","))
}

```


## 2. Data through 2022  

### 2.1 Raw data for analysis  

- We will also take basis into account  
- To simplify matters, we just modify PARAM so it also contains basis 
information  
- als

```{r}

# to recreate data from raw data, uncomment the following line:
# source("54data_Proref_2024.R")

data_all2 <- readRDS("Data/54_data_2024_loqfilter3x.rds")
df_series_sel <- readRDS("Data/54_dataseries_2024_loqfilter3x.rds")

cat("data_all2:", nrow(data_all2), "rows\n")
cat("df_series_sel:", nrow(df_series_sel), "rows\n")
cat("number of contaminants:", length(unique(df_series_sel$PARAM)), "\n")

# Contaminants to present in paper  
lookup_param <- readr::read_csv("Input_data/010E PARAMs list.txt")
cat("lookup_param:", nrow(lookup_param), "rows\n")

# What are the fate of "lost" contaminants - from script 54data_Proref_2024
params_fate <- readr::read_csv("Data/54_dataseries_2024_parameter_fate.csv")  

# How many of these 149 contaminants are covered in the data?
get_coverage(data_all2)  

# How many of these 149 contaminants are covered in the time series?
params_fate1 <- params_fate %>% 
  left_join(
    df_series_sel %>% 
      filter(grepl("__WW$", PARAM) & 
               Species %in% c("Gadus morhua", "Mytilus edulis")) %>% 
      distinct(PARAM) %>% 
      mutate(ts_orig = 1),
    by = join_by(PARAM_orig==PARAM)
  )

# Confirms that 139 are present
table(addNA(params_fate1$ts_orig))
 #   1 <NA> 
 # 139   10 

if (rerun_analysis == "test"){
  df_series_sel_original <- df_series_sel
}

```


### 2.2 LOQ levels and <LOQ frequency  

```{r}

years_plot <- 1992:2022

loq_summary <- subset(data_all2, YEAR %in% years_plot) %>%
  filter(Basis %in% "WW") %>%
  group_by(PARAM, LATIN_NAME, TISSUE_NAME, YEAR) %>%
  summarise(
    N = n(),
    N_under_LOQ = sum(FLAG1 %in% "<"),
    Perc_under_LOQ = N_under_LOQ/N*100,
    LOQ_median = median(Concentration[FLAG1 %in% "<"], na.rm = TRUE)
  ) 

# View(loq_summary)

# "CB_S7__WW", 
pars <- c("CB28__WW", "CB52__WW", "CB101__WW", "CB105__WW", "CB118__WW", "CB138__WW", 
"CB153__WW", "CB180__WW")

# data_all2 %>% filter(Basis %in% "WW" & PARAM %in% pars) %>% distinct(PARAM) %>% pull(PARAM) %>% dput()

ggplot(
  loq_summary %>% filter(PARAM %in% pars & LATIN_NAME == "Gadus morhua" &
                           TISSUE_NAME == "Lever"),
  aes(YEAR, LOQ_median, color = PARAM)) +
  geom_path() +
  scale_y_log10() +
  labs(title = "Cod, median LOQ")

ggplot(
  loq_summary %>% filter(PARAM %in% pars & LATIN_NAME == "Mytilus edulis" &
                           TISSUE_NAME == "Whole soft body"),
  aes(YEAR, LOQ_median, color = PARAM)) +
  geom_path() +
  scale_y_log10() +
  labs(title = "Blue mussel, median LOQ")

ggplot(
  loq_summary %>% filter(PARAM %in% pars & LATIN_NAME == "Gadus morhua" &
                           TISSUE_NAME == "Lever"),
  aes(YEAR, Perc_under_LOQ, color = PARAM)) +
  geom_path() +
  labs(title = "Cod, percent < LOQ")

ggplot(
  loq_summary %>% filter(PARAM %in% pars & LATIN_NAME == "Mytilus edulis" &
                           TISSUE_NAME == "Whole soft body"),
  aes(YEAR, Perc_under_LOQ, color = PARAM)) +
  geom_path() +
  labs(title = "Blue mussel, percent < LOQ")

```



## 3. Selection of background stations  

* For each contaminant / species  
    - Arrange stations from clean to more contaminated  
    - Stepwise testing (station 2 against station 1; station 3 against station 1+2; etc.)  
    - Stations are added to the list of background station until stopping  
* Creates three variables for three different stopping rules  
    - Background1 = TRUE for all stations not being different at P < 0.10   
    - Background1b = As Background1, but also the ratio beteween stations medians is not allowed to be above 2   
    - Background2 = TRUE for all stations not being different at P < 0.05 

### 3.1 Select background stations: test  

```{r}

# debugonce(get_background_values)

data_all3 <- data_all2[
  c("YEAR", "STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "SAMPLE_NO2", 
    "PARAM", "FLAG1", "Concentration", "UNIT")
]

# debugonce(get_background_values)
# debugonce(get_lower_medians)
test <- get_background_values(
       # determinant = "HG__WWa", 
       determinant = "BDE138__WW", 
       species = "Gadus morhua", 
       var_name = "Concentration", 
       years_backgr = years_backgr2,
       data = subset(data_all3, YEAR %in% years_backgr2))

# View(test$differences)     
# View(test$result_one_line)     

```

### 3.2 Select background stations: Run analysis for all data

* through 2022  
* Ca. 23 minutes with 4 paralell cores  
* NOTE: you can skip this by loading saved data - use code at the bottom of 3.4 (enclosed in "FALSE" brackets)  

```{r, results='hide'}

if (rerun_analysis == "test"){
  
  # test data: four series
  # including BDE138__WW which doesn't have enough data  
  df_series_sel <- df_series_sel_original %>% 
    filter(PARAM %in% c("HG__WW", "BDE138__WW"),
           Species %in% c("Gadus morhua", "Mytilus edulis"))
}

cat(hyphens, "Number of selected compounds x basis (per species)\n", hyphens)  
df_series_sel$Species %>% table()

# Decrease the number of columns to speed up parallel processing  
data_all3 <- data_all2[
  c("YEAR", "STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "SAMPLE_NO2", 
    "PARAM", "FLAG1", "Concentration", "UNIT")]

if (rerun_analysis == "no"){
  cat("saved detailed results: \n")
  dir("Data", "54_result_detailed_") %>% sort()
  cat("Read from file", sQuote(saved_detailed_results), "\n")
  result_detailed <- readRDS(saved_detailed_results)
  cat("number of contaminants:", nrow(result_detailed), "\n")
  cat("number of contaminants:", length(unique(result_detailed$PARAM)), "\n")

} else {
  
  # "Safe" version of function (doesn't break the loop if there are errors)
  get_background_values_s <- safely(get_background_values)
  
  plan(multisession, workers = 4)
  
  t0 <- Sys.time()
  # result_list is calculated using map along all rows of df_series_sel 
  # (not along all possible combinations of param and species)  
  result_list <- future_map(
    seq_len(nrow(df_series_sel)), 
    \(.x) get_background_values_s(
      determinant = df_series_sel$PARAM[.x], 
      species = df_series_sel$Species[.x], 
      var_name = "Concentration", 
      years_backgr = years_backgr2,
      data = subset(data_all3, YEAR %in% years_backgr2))
  )
  t1 <- Sys.time()
  t1-t0
  
}

# str(result_list[1:3], 2)

# writexl::write_xlsx(result, "Data/Proref_2022_version1.xlsx")   

```


#### Summary  

* Trenger  
    - hvor mange stasjoner man starta med  
    - varighet på hver stasjon  
    - hvor mange år (år x stasjon?) bygger proref på  

```{r}

if (rerun_analysis %in% c("test", "all")){
  
  cat("Time used:\n")
  t1-t0
  cat("\n")
  
  ok <- map(result_list, "error") %>% map_lgl(is.null)
  cat(hyphens, "Number of successful compound*species\n", hyphens)
  cat(sum(ok), "out of", length(ok), "\n")
  
  result <- map_dfr(result_list[ok], c("result", "result_one_line"))  
  cat(hyphens, "Result\n", hyphens)
  str(result)
  
}

```

### 3.3 Select background stations: detailed results  

This results in a data frame with one row for every step in the analysis (each station added) and three boolean (TRUE/FALSE) variables:    
- Background1 = TRUE for all stations not being different at P < 0.10  
- Background1b = As Background1, but also the ratio beteween stations medians is not allowed to be above 2  
- Background2 = TRUE for all stations not being different at P < 0.05       

```{r}

# str(result_list[[1]]$result, 1)
if (rerun_analysis %in% c("test", "all")){
  
  result_detailed <- map_dfr(result_list[ok], c("result", "differences"))  
  
  threshold_p1 <- 0.10
  threshold_p2 <- 0.05
  threshold_medianratio <- 2
  
  # Adding columns for background and last rank  
  
  # P > 0.1 limit:
  #   P_change1 = TRUE for the first station going over P > 0.1  
  #   P_sum1 = 0 for background stations, 1 for non-background  
  #   Background1 = background stations 
  #   Background1b = as Background1 but the ratio beteween stations medians is not allowed 
  #     to be above 2
  # P > 0.05 limit:
  #   P_change2 = TRUE for the first station going over P > 0.05  
  #   P_sum2 = 0 for background stations, 1 for non-background  
  
  
  result_detailed <- result_detailed %>% 
    # filter(PARAM %in% "CU" & LATIN_NAME %in% species) %>%   
    group_by(PARAM, LATIN_NAME, TISSUE_NAME) %>%
    mutate(
      P_round = round(P, 3),       # just for more easily readable output
      P_prev = lag(P_round),        # previous P value
      P_change1 = case_when(
        Rank == 1 ~ FALSE,
        TRUE ~ (P < threshold_p1) & (P_prev >= threshold_p1)), # change in P value
      P_cum1 = cumsum(P_change1), # cumulative change
      P_change2 = case_when(
        Rank == 1 ~ FALSE,
        TRUE ~ (P < threshold_p2) & (P_prev >= threshold_p2)), # change in P value
      P_cum2 = cumsum(P_change2), # cumulative change
      Background1 = case_when(
        P_cum1 == 0 ~ "Background",
        TRUE ~ "Other"),
      Background1b = case_when(
        P_cum1 == 0 & Median_ratio <= threshold_medianratio ~ "Background",
        TRUE ~ "Other"),
      Background2 = case_when(
        P_cum2 == 0 ~ "Background",
        TRUE ~ "Other"),
      Background = case_when(
        Background1b == "Background" ~ "Background 1b and 2",
        Background1 == "Background" ~ "Background 1 and 2",
        Background2 == "Background" ~ "Background 2",
        TRUE ~ "Other"),
      # This is used only for adding PARAM label to the last station
      #   in plots:
      Last_rank = (Rank == max(Rank))
    ) %>% 
    ungroup() %>%
    mutate(
      Line_label = case_when(
        Last_rank ~ PARAM,
        TRUE ~ NA)
    ) #%>%
  #select(-P_prev, -P_change1, -P_cum1, -P_change2, -P_cum2, -Last_rank)
  
  cat(hyphens, "Detailed result table\n", hyphens, sep = "")
  str(result_detailed)

}


```

### 3.4 Save and make table check 

```{r}

if (rerun_analysis == "all"){
  txt_years <- range(years_backgr2) %>% paste(collapse = "-")
  txt_timestamp <- lubridate::now() %>% substr(1, 16) %>% sub(" ", "-T", .) %>% sub(":", "", .)
  
  fn <- paste0("Data/54_result_detailed_", txt_years, "_", txt_timestamp, ".rds")
  saveRDS(result_detailed, fn) 
}

```

### 3.5 Plot detailed data - example    

```{r}

param <- "CU__WW"
species <- "Mytilus edulis"
param <- "HG__WWa"
species <- "Gadus morhua"

result_sel <- result_detailed %>%
  filter(PARAM == param & LATIN_NAME == species)

# Background stations - here using 'Background1b'  (HARD-CODED) 
result_sel %>%
  filter(Background1b == "Background") %>%
  pull(Station) %>%
  paste(collapse = ", ")

# Get medians (as a list)
annualmedians_result <- get_lower_medians(
  par = param, 
  sp = species, 
  ti = find_tissue(param, species, before_underscores = TRUE), 
  data = as.data.frame(subset(data_all2, YEAR %in% 1992:2022)), 
  variable_name = "Concentration", 
  min_indiv_per_year = find_min_indiv_per_year(species))
str(annualmedians_result, 1)
# annualmedians$median_per_station

ggplot(result_sel, aes(Rank, Median2)) +
  geom_line() +
  geom_point(aes(col = Background1b)) +
  geom_text_repel(aes(label = glue("{Station} (P = {P_round})")), hjust = 0, 
                  nudge_x = 0.5, point.padding = 0.2, box.padding = 0.25, direction = "y") +
  # Line at 2x the cleanest station
  geom_hline(aes(yintercept = 2*result_sel$Median2[1]), linetype = "dashed", colour = "brown") +
  expand_limits(y = 0)

ggplot(result_sel, aes(x = Min_year, xend = Max_year, y = Median2)) +
  geom_segment(aes(col = Background1b)) +
  expand_limits(y = 0)

# Combine medians to a data frame, and add the 'Background1b' column    
# Note: specifying 'Background1b' as the background we use here (HARD-CODED)
df_medians <- annualmedians_result$data_medians_list %>% 
  bind_rows() %>%
  rename(Station = STATION_CODE) %>%
  left_join(result_sel %>% select(Station, Background1b),
            by = join_by(Station), relationship = "many-to-one")

ggplot(df_medians, aes(YEAR, median, group = Station, colour = Background1b)) +
  geom_line()

ggplot(df_medians %>% filter(Background1b == "Background"), 
       aes(YEAR, median, colour = Station)) +
  geom_line() +
  geom_point()

```


### 3.6 "Lost" parameters  

#### "Lost" parameters 1    
```{r}

df_series_sel_summ <- df_series_sel %>% 
  filter(grepl("__WW$", PARAM) & 
           Species %in% c("Gadus morhua", "Mytilus edulis")) %>% 
  mutate(Species = factor(Species, levels = c("Gadus morhua", "Mytilus edulis"))) %>% 
  count(PARAM, Species, .drop = FALSE) %>%
  pivot_wider(names_from = Species, values_from = n, values_fill = 0, ) %>% 
  rename(ts_cod = `Gadus morhua`, ts_mussel = `Mytilus edulis`)

# 10 parameters were not in df_series_sel (ts == NA)
# no more parameters lost as number of stations were zero (ts == 0)
# 139 parameters still kept (ts == 1)
params_fate2 <- params_fate1 %>% 
  left_join(df_series_sel_summ, by = join_by(PARAM_orig == PARAM), relationship = "one-to-one") %>%
  mutate(ts = ifelse(ts_cod > 0 | ts_mussel > 0, 1, 0))
table(addNA(params_fate2$ts))

result_detailed_summ <- result_detailed %>% 
  filter(grepl("__WW$", PARAM) & 
           LATIN_NAME %in% c("Gadus morhua", "Mytilus edulis")) %>% 
  filter(Last_rank) %>% 
  mutate(LATIN_NAME = factor(LATIN_NAME, levels = c("Gadus morhua", "Mytilus edulis"))) %>% 
  count(PARAM, LATIN_NAME, .drop = FALSE) %>%
  pivot_wider(names_from = LATIN_NAME, values_from = n, values_fill = 0) %>% 
  rename(result1_cod = `Gadus morhua`, result1_mussel = `Mytilus edulis`)

nrow(result_detailed_summ)


# 30 more parameters lost
# ts was 1, result1 == NA
params_fate3 <- params_fate2 %>% 
  left_join(result_detailed_summ, by = join_by(PARAM_orig == PARAM), relationship = "one-to-one") %>%
  mutate(result1 = ifelse(result1_cod > 0 | result1_mussel > 0, 1, 0))
table(addNA(params_fate3$ts), addNA(params_fate3$result1))
  #        1 <NA>
  # 1    109   30
  # <NA>   0   10
params_lost3 <- params_fate3 %>% 
  filter(ts == 1 & is.na(result1)) %>% 
  pull(PARAM)
params_lost3
#  [1] "BDE138"             "BDE156"             "BDE191"             "BDE197"            
#  [5] "BDE205"             "BDE206"             "BDE207"             "PFHxS"             
#  [9] "Endrin"             "HCHB"               "Toksafen Parlar 62" "TTBT"              
# [13] "MOT"                "DOT"                "TCHT"               "MBT-Sn"            
# [17] "DBT-Sn"             "TBTIN"              "TTBT-Sn"            "TPhT-Sn"           
# [21] "MOT-Sn"             "TDCP"               "EHDPP"              "TEHP"              
# [25] "TCPP"               "4-N-OP"             "4-T-OP"             "4-N-NP"            
# [29] "4-T-NP"             "BPA" 

```

#### "Lost" parameters 2   
- These 30 parameters are lost because no stations have >= 5 years of data  
```{r}

subset(data_all3, YEAR %in% years_backgr2) %>% 
  filter(PARAM %in% "BDE138__WW") 

subset(data_all3, YEAR %in% years_backgr2) %>%
  ungroup() %>% 
  filter(
    PARAM %in% c(paste0(params_lost3, "__WW")),
    # PARAM %in% "BDE138__WW",
    LATIN_NAME %in% c("Gadus morhua", "Mytilus edulis")) %>%
  count(PARAM, LATIN_NAME, YEAR, STATION_CODE, name = "n_samples") %>% 
  mutate(
    min_n = 
      case_when(
        LATIN_NAME %in% "Gadus morhua" ~ 10,
        LATIN_NAME %in% "Mytilus edulis" ~ 2),
    year_accepted = n_samples > min_n) %>% 
  count(PARAM, LATIN_NAME, STATION_CODE, name = "n_years") %>%
  mutate(n_years_group = cut(n_years, breaks = c(0,2,3,4,5,15))) %>% 
  count(PARAM, LATIN_NAME, n_years_group) %>%
  pivot_wider(names_from = n_years_group, values_from = n, values_fill = 0)


```

## 4. Extract PROREF based on data from background stations   

* Proref is defined as the 95% percentile of concentrations in the background stations    
* This is implemented in the function get_conc_percentiles  

### 4.1. Reshape station data (defining background station)  

* reshaping 'result_detailed' to longer form  

```{r}

resultlist <- list(
    result_detailed %>% 
      filter(Background1b %in% "Background") %>%
      mutate(Background = "Background 1b"),
    result_detailed %>% 
      filter(Background1 %in% "Background") %>%
      mutate(Background = "Background 1"),
    result_detailed %>% 
      filter(Background2 %in% "Background") %>%
      mutate(Background = "Background 2"))

result_summ <- map_dfr(
  resultlist,
  \(df) df  %>%
    group_by(PARAM, LATIN_NAME) %>%
    summarize(
      Background = first(Background),
      Stations = paste(Station, collapse = ", "),
      Median_ratio = max(Median_ratio),
      n = n(), 
      Median_nyears = median(N_years),
      Median_firstyear = median(Min_year),
      Median_lastyear = median(Max_year),
      .groups = "drop"
    ))  %>%
  arrange(PARAM, LATIN_NAME, Background) 

```


### 4.2. Add potential number of stations  

```{r}

result_potential_number_of_stations <- result_detailed %>%
  count(PARAM, LATIN_NAME, name = "n_potential_stations")

result_summ <- result_summ %>%
  left_join(
    result_potential_number_of_stations, 
    by = join_by(PARAM, LATIN_NAME),
    relationship = "many-to-one"
  )

```

### 4.3 Add the actual PROREF values  

- Note: does not use 'get_conc_percentiles' function amymore - just using dplyr instead  
    - MUCH faster, takes only a second or two

- Background1 = TRUE for all stations not being different at P < 0.10  
- Background1b = As Background1, but also the ratio beteween stations medians is not allowed to be above 2 (the one used in the paper)     
- Background2 = TRUE for all stations not being different at P < 0.05      

```{r}

# Note: code is similar to code in script 55_plotting (part "raw data and proref value"), except
# - using 'Proref_type' instead of 'Analysis' (the latter refers to which 'result_detailed' file was used)
# - end result is called 'result_proref'  

# "lookup" data which tells (for each LATIN_NAME, PARAM) which stations that are background stations
lookup_background <- bind_rows(
  result_detailed %>%
    mutate(
      Proref_type = "p < 0.10, max median ratio",
      Background = Background1b) %>%
    select(Proref_type, LATIN_NAME, PARAM, Station, Background),
  result_detailed %>%
    mutate(
      Proref_type = "p < 0.10",
      Background = Background1) %>%
    select(Proref_type, LATIN_NAME, PARAM, Station, Background),
  result_detailed %>%
    mutate(
      Proref_type = "p < 0.05",
      Background = Background2) %>%
    select(Proref_type, LATIN_NAME, PARAM, Station, Background)
)

# xtabs(~Proref_type, lookup_background)
# xtabs(~Proref_type + Background, lookup_background)

# data for computing Proref: the raw data, then add info on whether the station
#   is a background station (for each LATIN_NAME, PARAM)
data_all_backgr <- data_all2 %>%
  ungroup() %>%
  select(Station = STATION_CODE, LATIN_NAME, PARAM, YEAR, Concentration, FLAG1) %>%
  # to the raw data (three lines above), we add the column Background1b (here called Background)
  # Station, LATIN_NAME, PARAM are used for joining data
  left_join(lookup_background, 
            by = join_by(Station, LATIN_NAME, PARAM), 
            relationship = "many-to-many") %>%
  # Keep only rows with some entry in "Background"
  filter(!is.na(Background)) %>%
  # Make new column 'Station_bg' which is the text entry 'Station' for each background station
  # (this is only used for plotting)
  mutate(
    Station_bg = case_when(
      Background %in% "Other" ~ NA,
      Background %in% "Background" ~ Station)
  ) %>%
  arrange(LATIN_NAME, PARAM, Proref_type, desc(Background), Station_bg) %>%
  # Define 'LOQ' column - also used only for plotting
  mutate(LOQ = case_when(
    is.na(FLAG1) ~ "Over LOQ",
    TRUE ~ "Under LOQ"))

#
# Proref
#
result_proref <- data_all_backgr %>% 
  filter(Background %in% "Background") %>%
  group_by(Proref_type, LATIN_NAME, PARAM) %>%
  summarize(PROREF = quantile(Concentration, 0.9) %>% signif(3),
            .groups = "drop")

```


#### Recreate the original PARAM and BASIS  

```{r}

# Split PARAM (e.g. 'PFOS__WW') up back into
# PARAM (e.g. 'PFOS') and BASIS (e.g. "WW)

X <- strsplit(result_proref$PARAM, split = "__")
result_proref$PARAM <- map_chr(X, 1)
result_proref$Basis <- map_chr(X, 2)
result_proref <- result_proref %>%
  select(PARAM, Basis, everything())

# View(result_proref)

```

### 4.4 Analysis of "lost" parameters  
```{r}

result_proref_wide <- result_proref %>% 
  # pivot_wider(names_from = Proref_type, values_from = PROREF)
  filter(
    Basis == "WW",
    Proref_type %in% "p < 0.10, max median ratio",
    LATIN_NAME %in% c("Gadus morhua", "Mytilus edulis")) %>% 
  select(PARAM, LATIN_NAME, PROREF) %>% 
  mutate(LATIN_NAME = factor(LATIN_NAME, levels = c("Gadus morhua", "Mytilus edulis"))) %>% 
  pivot_wider(names_from = LATIN_NAME, values_from = PROREF)

params_fate4 <- params_fate3 %>% 
  left_join(result_proref_wide, by = join_by(PARAM)) %>% 
  rename(
    proref_cod = `Gadus morhua`,
    proref_mussel = `Mytilus edulis`) %>% 
  mutate(
    proref = !(is.na(proref_cod) & is.na(proref_mussel))
  )

# 3 more parameters lost  
table(addNA(params_fate4$proref))
# FALSE  TRUE  <NA> 
#    43   106     0 

# Oarameters lost - 3 sums
params_fate4 %>% 
  filter(result1 == 1 & !proref) %>% 
  pull(PARAM)
# "LBsumPAH16" "LBsumBDE6"  "LBsumCB7"  

# Compare with parameters in paper
params_fate5 <- params_fate4 %>% 
  left_join(
    lookup_param %>% 
      rename(
        proref_cod_paper = `PROREF_Gadus morhua`,
        proref_mussel_paper = `PROREF_Mytilus edulis`) %>% 
      select(PARAM, proref_cod_paper, proref_mussel_paper)
  ) %>% 
  mutate(
    proref_paper = ifelse(is.na(proref_cod_paper) & is.na(proref_mussel_paper),0,1)
  )

# have lost 24 more
# now we have 82 
table(addNA(params_fate5$proref), addNA(params_fate5$proref_paper))

pars_lost <- params_fate5 %>% 
  filter(proref & is.na(proref_cod_paper) & is.na(proref_mussel_paper)) %>% 
  pull(PARAM)
pars_lost
#  [1] "BA"                    "MO"                    "V"                    
#  [4] "BDE85"                 "BDE184"                "BDE209"               
#  [7] "PFHxA"                 "PFHpA"                 "PFOA"                 
# [10] "PFNA"                  "PFBS"                  "Dieldrin"             
# [13] "alfa-Klordan (cis)"    "gamma-Klordan (trans)" "Oxyklordan"           
# [16] "Heptaklor epoksid"     "Nonaklor, trans-"      "Mirex"                
# [19] "Toksafen Parlar 26"    "Toksafen Parlar 50"    "DDTOP"                
# [22] "DDDOP"                 "DDEOP"                 "TBBPA"  

# these are lost because all the stations end up as bcaground stations
# (i.e., no station was significantly different from all stations )
result_detailed %>%
  filter(
    PARAM %in% c(paste0(pars_lost, "__WW"))) %>% 
  xtabs(~Background1b, .)

  

result_detailed %>%
  filter(
    PARAM %in% c(paste0(pars_lost, "__WW"))
    ) %>% 
  count(PARAM, Background1b) %>% 
  pivot_wider(names_from = Background1b, values_from = n)
  


```



## 5. Save  

* resu

```{r}

if (save_results){
  
  txt_years <- range(years_backgr2) %>% paste(collapse = "-")
  txt_timestamp <- lubridate::now() %>% substr(1, 16) %>% sub(" ", "-T", .) %>% sub(":", "", .)
  
  fn1 <- paste0("Data/54_result_summ_", txt_years, "_", txt_timestamp, ".csv")
  result_proref %>%
    readr::write_csv(fn1)
  cat("'result_proref' saved as ", fn1, "\n")
  
  fn2 <- paste0("Data/54_result_summ_final_", txt_years, "_", txt_timestamp, ".csv")
  result_proref %>%
    filter(Proref_type %in% "p < 0.10, max median ratio") %>%
    readr::write_csv(fn2)
  cat("'result_proref' with selected Proref_type saved as ", fn2, "\n")
  
}

```


## Appendix: Check and compare saved results  

```{r}

if (FALSE){
  
  # results calculated from code 
  result_proref_calc <- result_proref %>% filter(Proref_type %in% "p < 0.10, max median ratio")
  
  # results from saved file
  # Note: for results mad 2025-01-20 or earlier, the table contains very misleading values of 
  #   Prefer_type: "90 perc, max median ratio", "90 perc only", "95 perc"
  # This was change ini the code above 23.05.2025
  dir("Data", pattern = "54_result_summ_final_")
  result_proref_file <- readr::read_csv("Data/54_result_summ_final_2003-2022_2025-05-23-T1529.csv")
  
  names(result_proref_calc)
  names(result_proref_file)

  # compare
  library(compareDF)
  
  # for the entire table: only "Proref_type" differs
  ctable = compare_df(result_proref_calc, result_proref_file, c("PARAM", "Basis", "LATIN_NAME"))
  create_output_table(ctable)
    
  # for the entire table, and ignoring "Proref_type" 
  ctable = compare_df(result_proref_calc, result_proref_file, c("PARAM", "Basis", "LATIN_NAME"), 
                      exclude = "Proref_type", 
                      stop_on_error = FALSE)
  ctable$change_summary
  # the two data frames are the same  

}

```



